---
title: "Table2 Posteriors Bleeding"
author: "Jay Brophy"
output:
  html_document:
    code_folding: hide
    highlight: tango
    theme: cerulean
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
date: '`r format(Sys.Date(), "%B %d, %Y")`'
bibliography: [bib.bib]
biblio-style: apalike
---

```{r setup, include=FALSE, results='hide'}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 1, digits = 3)
library(tidyverse, brms) #bayesplot to install
library(brms)
library(bayesplot)
library(ggdist)
library(tidybayes)
library(ggplot2)
library(grid)
library(grDevices)
library(magrittr)
library(cmdstanr)


#tc4_dat1 <- readRDS("output/tc4_dat1.rds")
```

## Creating Table 2 Bleeding posterior probabilities

Priors in Table 2 include both vague and informative based on previous studies.

## Vague prior
### Posterior bleeding - vague prior with identity link
TC4 data not considering any clustering

```{r cache=TRUE, warning=FALSE, message=FALSE}

TC4_bleeding <- data.frame(y = c(28,20), N = c(555,450), drug = as.factor(c("0","1"))) # 0 = clopidogrel

b_prior_vg <- c(prior(student_t(3, log(1), 5 ), class=b, coef=drug0),
               prior(student_t(3, log(1), 5 ), class=b, coef=drug1)) # 0 = clopidogrel

f = bf(y | trials(N) ~ 0 + drug)

TC4_bleeding_vg <- brm(
  formula = f,
  data = TC4_bleeding,
  family = binomial(link = "identity"), #identity link probably better than get absolute numbers out can transform post
  chains = 4, warmup = 1000, iter = 2000, seed = 123,
  refresh = 0,
  prior=b_prior_vg, #prior(student_t(3, log(1), 5 ), class=b, coef=dapt_cat1)
  file = "TC4_bleeding_vg",
  file_refit = "on_change",
  refresh = 0
)


prior_summary(TC4_bleeding_vg)
print(TC4_bleeding_vg) # equivalent summary()

draws <- as_draws_df(TC4_bleeding_vg) # posterior_samples is equivalent outdated command

draws <- draws %>% 
  # rename and drop the unneeded columns
  transmute(p0 = b_drug0,   # expit not used since identity link
            p1 = b_drug1) %>% 
  # compute the OR
  mutate(or = (p1 / (1 - p1)) / (p0 / (1 - p0)), rr = p1 /p0)


temp <- round(quantile(draws$rr, c(0.025, 0.5, 0.975)),2)
paste0("RR = ", temp[2],  " 95% CI ",  temp[1], " - ", temp[3] )


# bleeding model - vg prior
draws %>% 
  ggplot(aes(x=rr, y=0)) + 
  stat_halfeye(.width = c(0.025, 0.975)) + 
  scale_y_continuous(NULL, breaks = NULL) + 
  xlab("Strok RR (ticagrelor vs. clopidogrel[ref])") +
  theme(panel.grid = element_blank()) +
  geom_vline(xintercept = c(0.9,1.11), colour="black", linetype="dashed") +
  geom_vline(xintercept = 1, colour="black", linetype="solid") +
  scale_x_continuous(limits = c(0, 2), expand = c(0, 0)) +
  ggtitle("Bleeding with vague prior")


plot(TC4_bleeding_vg)
grDevices::cairo_pdf("output/TC4_bleeding_vg.pdf", width = 6, height = 6, family = "Arial")
plot(TC4_bleeding_vg, digits = 2, base_size = 10)
dev.off()

prop0.9_mi <- sum(draws$rr<0.9)/nrow(draws)
prop1.0_mi <- sum(draws$rr<1.0)/nrow(draws)
prop1.1_mi <- sum(draws$rr>1.1)/nrow(draws)

cat("Pr(HR<0.9): ", prop0.9_mi,
    "\nPr(HR<1.0): ", prop1.0_mi,
    "\nPr(HR>1.1): ", prop1.1_mi,
    "\nPr(HR[0.9, 1.1]): ", 1-prop1.1_mi-prop0.9_mi)

```
Using a vague deaths prior , the MI posterior distribution is a N(`r temp[2]`, `r (temp[3] - temp[2])/1.96 `)     


Diagnostics
```{r}
pairs(TC4_bleeding_vg)
plot(TC4_bleeding_vg, ask=F)
```

```{r}
mcmc_areas(draws,
           pars = "rr",
           prob = 0.95)
```

## PLATO US bleeding prior 

US PLATO bleeding data available from Mahaffey[@PLATO_geo]

Use `brms` to get the posterior distribution for the PLATO US stroke prior

```{r cache=TRUE, warning=FALSE, message=FALSE, results='hide'}
set.seed(123)

TC4_bleeding_NA <- brm(
  formula = f,
  data = TC4_bleeding,
  family = binomial(link = "identity"), #identity link probably better than get absolute numbers out can transform post
  chains = 4, warmup = 1000, iter = 2000, seed = 123,
  refresh = 0,
  prior = c(prior(beta(74,601), class = b, coef = "drug0"),  
            prior(beta(77,605),class = b, coef = "drug1")), 
#  file = "TC4_bleeding_NA",
#  file_refit = "on_change",
  refresh = 0
)
```

```{r cache=TRUE, warning=FALSE, message=FALSE}
prior_summary(TC4_bleeding_NA)
print(TC4_bleeding_NA) # equivalent summary(m_logit)

draws <- as_draws_df(TC4_bleeding_NA) # posterior_samples is equivalent command

draws <- draws %>% 
  # rename and drop the unneeded columns
  transmute(p0 = b_drug0,   # expit not used since identity link
            p1 = b_drug1) %>% 
  # compute the OR
  mutate(or = (p1 / (1 - p1)) / (p0 / (1 - p0)), rr = p1 /p0)

temp <- round(quantile(draws$rr, c(0.025, 0.5, 0.975)),2)
paste0("RR = ", temp[2],  " 95% CI ",  temp[1], " - ", temp[3] )


# bleeding model - PLATO prior
draws %>% 
  ggplot(aes(x=rr, y=0)) + 
  stat_halfeye(.width = c(0.025, 0.975)) + 
  scale_y_continuous(NULL, breaks = NULL) + 
  xlab("Bleeding RR (ticagrelor vs. clopidogrel[ref])") +
  theme(panel.grid = element_blank()) +
  geom_vline(xintercept = c(0.9,1.11), colour="black", linetype="dashed") +
  geom_vline(xintercept = 1, colour="black", linetype="solid") +
  scale_x_continuous(limits = c(0.5, 2), expand = c(0, 0)) +
  ggtitle("Major bleeding with US prior")


plot(TC4_bleeding_NA)
grDevices::cairo_pdf("output/bleeding_NA.pdf", width = 6, height = 6, family = "Arial")
plot(TC4_bleeding_NA, digits = 2, base_size = 10)
dev.off()

prop0.9_mi <- sum(draws$rr<0.9)/nrow(draws)
prop1.0_mi <- sum(draws$rr<1.0)/nrow(draws)
prop1.1_mi <- sum(draws$rr>1.1)/nrow(draws)

cat("Pr(HR<0.9): ", prop0.9_mi,
    "\nPr(HR<1.0): ", prop1.0_mi,
    "\nPr(HR>1.1): ", prop1.1_mi,
    "\nPr(HR[0.9, 1.1]): ", 1-prop1.1_mi-prop0.9_mi)

```

## PLATO all data bleeding prior 

Use `brms` to get the posterior distribution for the PLATO bleeding prior

```{r cache=TRUE, warning=FALSE, message=FALSE, results='hide'}

TC4_bleeding_PLATO <- brm(
  formula = f,
  data = TC4_bleeding,
  family = binomial(link = "identity"), #identity link probably better than get absolute numbers out can transform post
  chains = 4, warmup = 1000, iter = 2000, seed = 123,
  refresh = 0,
  prior = c(prior(beta(929,8257), class = b, coef = "drug0"),  
            prior(beta(961,8274),class = b, coef = "drug1")), 
#  file = "TC4_bleeding_PLATO",
#  file_refit = "on_change",
  refresh = 0
)
```

```{r cache=TRUE, warning=FALSE, message=FALSE}
prior_summary(TC4_bleeding_PLATO)
print(TC4_bleeding_PLATO) # equivalent summary(m_logit)

draws <- as_draws_df(TC4_bleeding_PLATO) # posterior_samples is equivalent command

draws <- draws %>% 
  # rename and drop the unneeded columns
  transmute(p0 = b_drug0,   # expit not used since identity link
            p1 = b_drug1) %>% 
  # compute the OR
  mutate(or = (p1 / (1 - p1)) / (p0 / (1 - p0)), rr = p1 /p0)

temp <- round(quantile(draws$rr, c(0.025, 0.5, 0.975)),2)
paste0("RR = ", temp[2],  " 95% CI ",  temp[1], " - ", temp[3] )


# bleeding model - PLATO prior
draws %>% 
  ggplot(aes(x=rr, y=0)) + 
  stat_halfeye(.width = c(0.025, 0.975)) + 
  scale_y_continuous(NULL, breaks = NULL) + 
  xlab("bleeding RR (ticagrelor vs. clopidogrel[ref])") +
  theme(panel.grid = element_blank()) +
  geom_vline(xintercept = c(0.9,1.11), colour="black", linetype="dashed") +
  geom_vline(xintercept = 1, colour="black", linetype="solid") +
  scale_x_continuous(limits = c(0.5, 2), expand = c(0, 0)) +
  ggtitle("bleeding with PLATO prior")


plot(TC4_bleeding_PLATO)
grDevices::cairo_pdf("output/bleeding_PLATO.pdf", width = 6, height = 6, family = "Arial")
plot(TC4_bleeding_PLATO, digits = 2, base_size = 10)
dev.off()

prop0.9_mi <- sum(draws$rr<0.9)/nrow(draws)
prop1.0_mi <- sum(draws$rr<1.0)/nrow(draws)
prop1.1_mi <- sum(draws$rr>1.1)/nrow(draws)

cat("Pr(HR<0.9): ", prop0.9_mi,
    "\nPr(HR<1.0): ", prop1.0_mi,
    "\nPr(HR>1.1): ", prop1.1_mi,
    "\nPr(HR[0.9, 1.1]): ", 1-prop1.1_mi-prop0.9_mi)

```

Using a full PLATO bleeding prior, the  bleeding posterior distribution is a N(`r temp[2]`, `r (temp[3] - temp[2])/1.96 `)       

Diagnostics
```{r}
pairs(TC4_bleeding_PLATO)
plot(TC4_bleeding_PLATO, ask=F)
```

```{r}
mcmc_areas(draws,
           pars = "rr",
           prob = 0.95)
```


## BNMA bleeding prior 

Use BNMA data to create priors for effectiveness of ticagrelor  (203 / 15373) vs. clopidogrel (305 / 25543), which is  N(1.16, 0.16) as the prior for the BNMA data


Use `brms` to get the posterior distribution for the BNMA bleeding prior

```{r cache=TRUE, warning=FALSE, message=FALSE, results='hide'}

TC4_bleeding_BNMA <- brm(
  formula = f,
  data = TC4_bleeding,
  family = binomial(link = "identity"), #identity link probably better than get absolute numbers out can transform post
  chains = 4, warmup = 1000, iter = 2000, seed = 123,
  refresh = 0,
#  prior = c(prior(normal(1, .05), class = b, coef ="drug0"),
#            prior(normal(1.1, 0.06), class = b, coef ="drug1")),
  prior = c(prior(beta(1030,12149), class = b, coef = "drug0"),  
           prior(beta(1113,12145),class = b, coef = "drug1")), 
#  file = "TC4_bleeding_PLATO_brms",
#  file_refit = "on_change",
  refresh = 0
)
```

```{r cache=TRUE, warning=FALSE, message=FALSE}
prior_summary(TC4_bleeding_BNMA)
print(TC4_bleeding_BNMA) # equivalent summary(m_logit)

draws <- as_draws_df(TC4_bleeding_BNMA) # posterior_samples is equivalent command

draws <- draws %>% 
  # rename and drop the unneeded columns
  transmute(p0 = b_drug0,   # expit not used since identity link
            p1 = b_drug1) %>% 
  # compute the OR
  mutate(or = (p1 / (1 - p1)) / (p0 / (1 - p0)), rr = p1 /p0)

temp <- round(quantile(draws$rr, c(0.025, 0.5, 0.975)),2)
paste0("RR = ", temp[2],  " 95% CI ",  temp[1], " - ", temp[3] )


# bleeding model - BNMA prior
draws %>% 
  ggplot(aes(x=rr, y=0)) + 
  stat_halfeye(.width = c(0.025, 0.975)) + 
  scale_y_continuous(NULL, breaks = NULL) + 
  xlab("bleeding RR (ticagrelor vs. clopidogrel[ref])") +
  theme(panel.grid = element_blank()) +
  geom_vline(xintercept = c(0.9,1.11), colour="black", linetype="dashed") +
  geom_vline(xintercept = 1, colour="black", linetype="solid") +
  scale_x_continuous(limits = c(0.5, 2), expand = c(0, 0)) +
  ggtitle("BNMA prior")


plot(TC4_bleeding_BNMA)
grDevices::cairo_pdf("output/bleeding_BNMA.pdf", width = 6, height = 6, family = "Arial")
plot(TC4_bleeding_BNMA, digits = 2, base_size = 10)
dev.off()

prop0.9_mi <- sum(draws$rr<0.9)/nrow(draws)
prop1.0_mi <- sum(draws$rr<1.0)/nrow(draws)
prop1.1_mi <- sum(draws$rr>1.1)/nrow(draws)

cat("Pr(HR<0.9): ", prop0.9_mi,
    "\nPr(HR<1.0): ", prop1.0_mi,
    "\nPr(HR>1.1): ", prop1.1_mi,
    "\nPr(HR[0.9, 1.1]): ", 1-prop1.1_mi-prop0.9_mi)

```

Using a full PLATO bleeding prior, the  bleeding posterior distribution is a N(`r temp[2]`, `r (temp[3] - temp[2])/1.96 `)       

Diagnostics
```{r}
pairs(TC4_bleeding_BNMA)
plot(TC4_bleeding_BNMA, ask=F)
```

```{r}
mcmc_areas(draws,
           pars = "rr",
           prob = 0.95)
```
## References