---
title: "Table  MACE"
author: "Jay Brophy"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: hide
    highlight: tango
    theme: cerulean
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
link-citations: yes
linkcolor: red
urlcolor: blue
bibliography: [bib.bib]
biblio-style: apalike
---

```{r setup, echo=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.align = "center",
  fig.asp = 0.618,
  fig.retina = 3,
  fig.width = 6,
  message = FALSE,
  warning = FALSE,
  dev = "svg",
  out.width = "80%")

library(magrittr)
library(dplyr)
library(purrr)
library(forcats)
library(tidyr)
library(modelr)
library(ggdist)
library(tidybayes)
library(ggplot2)
library(cowplot)
library(rstan)
library(rstanarm)
library(brms)
library(cmdstanr)
library(RColorBrewer)
library(here)
library(knitr)
library(survey)
library(tidyverse)
library(broom)
library(ggdag)
library(baymedr) # devtools::install_github("maxlinde/baymedr")
library(gt) 
options(knitr.table.format = "html")
options(knitr.kable.NA = '   ')
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
options(scipen = 1, digits = 3) #set to two decimal 

```

## MACE

### Vague prior

```{r results='hide', cache=TRUE}

data_TC4 <- data.frame(N = c(550,450), y = c(64,50), drug = as.factor(c(0,1))) # 0 = clopidogrel

f = bf(y | trials(N) ~ 0 + drug)

m1 <- brm(
  formula = f,
  data = data_TC4,
  family = binomial(link = "identity"),
  chains = 4, warmup = 1000, iter = 5000, seed = 123,
  refresh = 0
)


```


```{r}

prior_summary(m1)
print(m1) # equivalent summary()

draws <- as_draws_df(m1) # posterior_samples is equivalent outdated command

draws <- draws %>% 
  # rename and drop the unneeded columns
  transmute(p0 = b_drug0,   # expit not used since identity link
            p1 = b_drug1) %>% 
  # compute the OR
  mutate(or = (p1 / (1 - p1)) / (p0 / (1 - p0)), rr = p1 /p0)


temp <- round(quantile(draws$rr, c(0.025, 0.5, 0.975)),2)
paste0("RR = ", temp[2],  " 95% CI ",  temp[1], " - ", temp[3] )


# bleeding model - vg prior
draws %>% 
  ggplot(aes(x=rr, y=0)) + 
  stat_halfeye(.width = c(0.025, 0.975)) + 
  scale_y_continuous(NULL, breaks = NULL) + 
  xlab("MACE RR (ticagrelor vs. clopidogrel[ref])") +
  theme(panel.grid = element_blank()) +
  geom_vline(xintercept = c(0.9,1.11), colour="black", linetype="dashed") +
  geom_vline(xintercept = 1, colour="black", linetype="solid") +
  scale_x_continuous(limits = c(0, 2), expand = c(0, 0)) +
  ggtitle("MACE with vague prior")


plot(m1)
grDevices::cairo_pdf("output/mace_vg.pdf", width = 6, height = 6, family = "Arial")
plot(m1, digits = 2, base_size = 10)
dev.off()

prop0.9 <- sum(draws$rr<0.9)/nrow(draws)
prop1.0 <- sum(draws$rr<1.0)/nrow(draws)
prop1.1 <- sum(draws$rr>1.1)/nrow(draws)

cat("Pr(HR<0.9): ", prop0.9,
    "\nPr(HR<1.0): ", prop1.0,
    "\nPr(HR>1.1): ", prop1.1,
    "\nPr(HR[0.9, 1.1]): ", 1-prop1.1-prop0.9)

```

### Informative prior - NA PLATO

Data from here[@PLATO_geo]
US data from PLATO trial ticagrelor (89/706) vs. clopidogrel (71/707)
CDN data from PLATO trial ticagrelor (18/201) vs. clopidogrel (15/200)


```{r results='hide', cache=TRUE}
# PLATO NA

m1.NA <- brm(
  formula = f,
  data = data_TC4,
  family = binomial(link = "identity"),
  prior = c(prior(beta(107,800), class = b, coef = "drug1"),
            prior(beta(86,821),class = b, coef = "drug0")),
  chains = 4, warmup = 1000, iter = 5000, seed = 123,
  refresh = 0
)
```


```{r}
prior_summary(m1.NA)
print(m1.NA) # equivalent summary()

draws <- as_draws_df(m1.NA) # posterior_samples is equivalent outdated command

draws <- draws %>% 
  # rename and drop the unneeded columns
  transmute(p0 = b_drug0,   # expit not used since identity link
            p1 = b_drug1) %>% 
  # compute the OR
  mutate(or = (p1 / (1 - p1)) / (p0 / (1 - p0)), rr = p1 /p0)


temp <- round(quantile(draws$rr, c(0.025, 0.5, 0.975)),2)
paste0("RR = ", temp[2],  " 95% CI ",  temp[1], " - ", temp[3] )


# MACE model - NA prior
draws %>% 
  ggplot(aes(x=rr, y=0)) + 
  stat_halfeye(.width = c(0.025, 0.975)) + 
  scale_y_continuous(NULL, breaks = NULL) + 
  xlab("MACE RR (ticagrelor vs. clopidogrel[ref])") +
  theme(panel.grid = element_blank()) +
  geom_vline(xintercept = c(0.9,1.11), colour="black", linetype="dashed") +
  geom_vline(xintercept = 1, colour="black", linetype="solid") +
  scale_x_continuous(limits = c(0, 2), expand = c(0, 0)) +
  ggtitle("MACE with PLATO NA only data prior")


plot(m1.NA)
grDevices::cairo_pdf("output/mace_na.pdf", width = 6, height = 6, family = "Arial")
plot(m1.NA, digits = 2, base_size = 10)
dev.off()

prop0.9 <- sum(draws$rr<0.9)/nrow(draws)
prop1.0 <- sum(draws$rr<1.0)/nrow(draws)
prop1.1 <- sum(draws$rr>1.1)/nrow(draws)

cat("Pr(HR<0.9): ", prop0.9,
    "\nPr(HR<1.0): ", prop1.0,
    "\nPr(HR>1.1): ", prop1.1,
    "\nPr(HR[0.9, 1.1]): ", 1-prop1.1-prop0.9)

```

### Informative prior - All PLATO

Prior assuming no between country heterogeneity
```{r results='hide', cache=TRUE}
# All PLATO

m1.PLATO <- brm(
  formula = f,
  data = data_TC4,
  family = binomial(link = "identity"),
  prior = c(prior(beta(864,8469), class = b, coef = "drug1"),
            prior(beta(1014,8277),class = b, coef = "drug0")),
  chains = 4, warmup = 1000, iter = 5000, seed = 123,
  refresh = 0
)
```


```{r}
prior_summary(m1.PLATO)
print(m1.PLATO) # equivalent summary()

draws <- as_draws_df(m1.PLATO) # posterior_samples is equivalent outdated command

draws <- draws %>% 
  # rename and drop the unneeded columns
  transmute(p0 = b_drug0,   # expit not used since identity link
            p1 = b_drug1) %>% 
  # compute the OR
  mutate(or = (p1 / (1 - p1)) / (p0 / (1 - p0)), rr = p1 /p0)


temp <- round(quantile(draws$rr, c(0.025, 0.5, 0.975)),2)
paste0("RR = ", temp[2],  " 95% CI ",  temp[1], " - ", temp[3] )


# MACE model - PLATO prior
draws %>% 
  ggplot(aes(x=rr, y=0)) + 
  stat_halfeye(.width = c(0.025, 0.975)) + 
  scale_y_continuous(NULL, breaks = NULL) + 
  xlab("MACE RR (ticagrelor vs. clopidogrel[ref])") +
  theme(panel.grid = element_blank()) +
  geom_vline(xintercept = c(0.9,1.11), colour="black", linetype="dashed") +
  geom_vline(xintercept = 1, colour="black", linetype="solid") +
  scale_x_continuous(limits = c(0, 2), expand = c(0, 0)) +
  ggtitle("MACE with all PLATO data prior")


plot(m1.PLATO)
grDevices::cairo_pdf("output/mace_PLATO.pdf", width = 6, height = 6, family = "Arial")
plot(m1.PLATO, digits = 2, base_size = 10)
dev.off()

prop0.9 <- sum(draws$rr<0.9)/nrow(draws)
prop1.0 <- sum(draws$rr<1.0)/nrow(draws)
prop1.1 <- sum(draws$rr>1.1)/nrow(draws)

cat("Pr(HR<0.9): ", prop0.9,
    "\nPr(HR<1.0): ", prop1.0,
    "\nPr(HR>1.1): ", prop1.1,
    "\nPr(HR[0.9, 1.1]): ", 1-prop1.1-prop0.9)

```

However this prior ignores the heterogenity within the PLATO data[@PLATO] according to country.   

```{r fig.height=12}
knitr::include_graphics("Country_mace_MA2.pdf")
```

A more realistic PLATO prior would include this variability. The point estimate remains the same but would use standard deviation of the prediction interval for the geo-stratified next study. In this case the prior would be N(0.84, 0.14) [@Kutcher].

```{r results='hide', cache=TRUE}
# All PLATO

m2.PLATO <- brm(
  formula = f,
  data = data_TC4,
  family = binomial(link = "identity"),
  prior = prior(normal(0.84,0.14), class = b, coef = "drug1"),
  chains = 4, warmup = 1000, iter = 5000, seed = 123,
  refresh = 0
)
```


```{r}
prior_summary(m2.PLATO)
print(m2.PLATO) # equivalent summary()

draws <- as_draws_df(m2.PLATO) # posterior_samples is equivalent outdated command

draws <- draws %>% 
  # rename and drop the unneeded columns
  transmute(p0 = b_drug0,   # expit not used since identity link
            p1 = b_drug1) %>% 
  # compute the OR
  mutate(or = (p1 / (1 - p1)) / (p0 / (1 - p0)), rr = p1 /p0)


temp <- round(quantile(draws$rr, c(0.025, 0.5, 0.975)),2)
paste0("RR = ", temp[2],  " 95% CI ",  temp[1], " - ", temp[3] )


# MACE model - PLATO prior accounting for between country heterogeneity
draws %>% 
  ggplot(aes(x=rr, y=0)) + 
  stat_halfeye(.width = c(0.025, 0.975)) + 
  scale_y_continuous(NULL, breaks = NULL) + 
  xlab("MACE RR (ticagrelor vs. clopidogrel[ref])") +
  theme(panel.grid = element_blank()) +
  geom_vline(xintercept = c(0.9,1.11), colour="black", linetype="dashed") +
  geom_vline(xintercept = 1, colour="black", linetype="solid") +
  scale_x_continuous(limits = c(0, 2), expand = c(0, 0)) +
  ggtitle("MACE with all PLATO data prior")


plot(m1.PLATO)
grDevices::cairo_pdf("output/mace_PLATO.pdf", width = 6, height = 6, family = "Arial")
plot(m1.PLATO, digits = 2, base_size = 10)
dev.off()

prop0.9 <- sum(draws$rr<0.9)/nrow(draws)
prop1.0 <- sum(draws$rr<1.0)/nrow(draws)
prop1.1 <- sum(draws$rr>1.1)/nrow(draws)

cat("Pr(HR<0.9): ", prop0.9,
    "\nPr(HR<1.0): ", prop1.0,
    "\nPr(HR>1.1): ", prop1.1,
    "\nPr(HR[0.9, 1.1]): ", 1-prop1.1-prop0.9)

```

### Informative prior - BNMA

Data from Bayesian network meta-analysisprediction interval for the next study[@Kutcher]

```{r results='hide', cache=TRUE}
# BNMA

m1.BNMA <- brm(
  formula = f,
  data = data_TC4,
  family = binomial(link = "identity"),
  prior = prior(normal(1.0,.25), class = b, coef = "drug1"), #from PI_mace.Rmd
  chains = 4, warmup = 1000, iter = 5000, seed = 123,
  refresh = 0
)


```


```{r}
prior_summary(m1.BNMA)
print(m1.BNMA) # equivalent summary()

draws <- as_draws_df(m1.BNMA) # posterior_samples is equivalent outdated command

draws <- draws %>% 
  # rename and drop the unneeded columns
  transmute(p0 = b_drug0,   # expit not used since identity link
            p1 = b_drug1) %>% 
  # compute the OR
  mutate(or = (p1 / (1 - p1)) / (p0 / (1 - p0)), rr = p1 /p0)


temp <- round(quantile(draws$rr, c(0.025, 0.5, 0.975)),2)
paste0("RR = ", temp[2],  " 95% CI ",  temp[1], " - ", temp[3] )


# MACE model - PLATO prior
draws %>% 
  ggplot(aes(x=rr, y=0)) + 
  stat_halfeye(.width = c(0.025, 0.975)) + 
  scale_y_continuous(NULL, breaks = NULL) + 
  xlab("MACE RR (ticagrelor vs. clopidogrel[ref])") +
  theme(panel.grid = element_blank()) +
  geom_vline(xintercept = c(0.9,1.11), colour="black", linetype="dashed") +
  geom_vline(xintercept = 1, colour="black", linetype="solid") +
  scale_x_continuous(limits = c(0, 2), expand = c(0, 0)) +
  ggtitle("MACE with all PLATO data prior")


plot(m1.BNMA)
grDevices::cairo_pdf("output/mace_BNMA.pdf", width = 6, height = 6, family = "Arial")
plot(m1.BNMA, digits = 2, base_size = 10)
dev.off()

prop0.9 <- sum(draws$rr<0.9)/nrow(draws)
prop1.0 <- sum(draws$rr<1.0)/nrow(draws)
prop1.1 <- sum(draws$rr>1.1)/nrow(draws)

cat("Pr(HR<0.9): ", prop0.9,
    "\nPr(HR<1.0): ", prop1.0,
    "\nPr(HR>1.1): ", prop1.1,
    "\nPr(HR[0.9, 1.1]): ", 1-prop1.1-prop0.9)

```