---
title: "Table2 Posteriors MI"
author: "Jay Brophy"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: hide
    highlight: tango
    theme: cerulean
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
link-citations: yes
linkcolor: red
urlcolor: blue
bibliography: [bib.bib]
biblio-style: apalike
---

```{r setup, echo=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.align = "center")

library(magrittr)
library(dplyr)
library(purrr)
library(forcats)
library(tidyr)
library(modelr)
library(ggdist)
library(tidybayes)
library(ggplot2)
library(cowplot)
library(rstan)
library(rstanarm)
library(brms)
library(cmdstanr)
library(RColorBrewer)
library(here)
library(knitr)
library(survey)
library(tidyverse)
library(broom)
library(ggdag)
library(baymedr) # devtools::install_github("maxlinde/baymedr")
library(gt) 
options(knitr.table.format = "html")
options(knitr.kable.NA = '   ')
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
options(scipen = 1, digits = 3) #set to two decimal 

```


## MI posterior probabilities
### Vague prior
TC4 data not considering any time clustering.   
This is done using both logit and identity links

## Possible Priors
```{r}

b_prior_NA <- c(prior(normal(.16,.1), class=b, coef=drug0),
              prior(normal(.195,.1), class=b, coef=drug1))
# b_prior_NA <- prior(normal(log(1.23), ((log(1.50)-log(1.01))/3.92)), class=b, coef=dapt) # PLATO NA prior from US & Canada data above
b_prior_plato <- prior(normal(log(0.79), ((log(1.15)-log(0.55))/3.92)), class=b, coef=dapt)  # hierarchical PI
b_prior_plato_pooled <- prior(normal(log(0.84), ((log(0.92)-log(0.77))/3.92)), class=b, coef=dapt) # NEJM pooled analysis
b_prior_bnma <- prior(normal(log(0.94), ((log(1.17)-log(0.76))/3.92)), class=b, coef=dapt) # stephen thesis supplement table

```

Posterior MI - vague prior with logit link


```{r cache=TRUE, warning=FALSE, message=FALSE}

TC4_mi <- data.frame(y = c(46,38), N = c(555,450), drug = as.factor(c("0","1"))) # 0 = clopidogrel
f = bf(y | trials(N) ~ 0 + drug)
b_prior_vg <- c(prior(student_t(3, log(1), 5 ), class=b, coef=drug0),
               prior(student_t(3, log(1), 5 ), class=b, coef=drug1)) # 0 = clopidogrel

tc4_mi_vg <- brm(
  formula = f,
  data = TC4_mi,
  family = binomial(link = "logit"), #identity link probably better than get absolute numbers out can transform post
  chains = 4, warmup = 1000, iter = 2000, seed = 123,
  refresh = 0,
  prior=b_prior_vg, #prior(student_t(3, log(1), 5 ), class=b, coef=dapt_cat1)
  file = "tc4_bayes_mi_vg_binomial",
  refresh = 0
)
```

```{r}
prior_summary(tc4_mi_vg)
print(tc4_mi_vg) # equivalent summary()

draws <- as_draws_df(tc4_mi_vg) # posterior_samples is equivalent outdated command

# since working with binomial link need this function to back transform
expit <- function(x){
  return(exp(x)/(1+exp(x)))
}

draws <- draws %>% 
  # rename and drop the unneeded columns
  transmute(p0 = expit(b_drug0),        # expit since used logit link no need for this function if identity link
            p1 = expit(b_drug1)) %>% 
  # compute the OR
  mutate(or = (p1 / (1 - p1)) / (p0 / (1 - p0)), rr = p1 /p0)

temp <- round(quantile(draws$rr, c(0.025, 0.5, 0.975)),2)
paste0("RR = ", temp[2],  " 95% CI ",  temp[1], " - ", temp[3] )


# MI model - vg prior
draws %>% 
  ggplot(aes(x=rr, y=0)) + 
  stat_halfeye(.width = c(0.025, 0.975)) + 
  scale_y_continuous(NULL, breaks = NULL) + 
  xlab("MI RR (ticagrelor vs. clopidogrel[ref])") +
  theme(panel.grid = element_blank()) +
  geom_vline(xintercept = c(0.9,1.11), colour="black", linetype="dashed") +
  geom_vline(xintercept = 1, colour="black", linetype="solid") +
  scale_x_continuous(limits = c(0, 2), expand = c(0, 0)) +
  ggtitle("MI with vague prior")


plot(tc4_mi_vg)
grDevices::cairo_pdf("output/MI_vg.pdf", width = 6, height = 6, family = "Arial")
plot(tc4_mi_vg, digits = 2, base_size = 10)
dev.off()

prop0.9_mi <- sum(draws$rr<0.9)/nrow(draws)
prop1.0_mi <- sum(draws$rr<1.0)/nrow(draws)
prop1.1_mi <- sum(draws$rr>1.1)/nrow(draws)

cat("Pr(HR<0.9): ", prop0.9_mi,
    "\nPr(HR<1.0): ", prop1.0_mi,
    "\nPr(HR>1.1): ", prop1.1_mi,
    "\nPr(HR[0.9, 1.1]): ", 1-prop1.1_mi-prop0.9_mi)

```
Using a vague deaths prior , the MI posterior distribution is a N(`r temp[2]`, `r (temp[3] - temp[2])/1.96 `)     


Posterior MI - vague prior with identity link

```{r cache=TRUE, warning=FALSE, message=FALSE}

tc4_mi_vg_identity <- brm(
  formula = f,
  data = TC4_mi,
  family = binomial(link = "identity"), #identity link probably better than get absolute numbers out can transform post
  chains = 4, warmup = 1000, iter = 2000, seed = 123,
  refresh = 0,
  prior=b_prior_vg, #prior(student_t(3, log(1), 5 ), class=b, coef=dapt_cat1)
  refresh = 0
)

```

```{r}
prior_summary(tc4_mi_vg_identity)
print(tc4_mi_vg_identity) # equivalent summary(m_logit)

draws <- as_draws_df(tc4_mi_vg_identity) # posterior_samples is equivalent command

draws <- draws %>% 
  # rename and drop the unneeded columns
  transmute(p0 = b_drug0,   # expit not used since identity link
            p1 = b_drug1) %>% 
  # compute the OR
  mutate(or = (p1 / (1 - p1)) / (p0 / (1 - p0)), rr = p1 /p0)


temp <- round(quantile(draws$rr, c(0.025, 0.5, 0.975)),2)
paste0("RR = ", temp[2],  " 95% CI ",  temp[1], " - ", temp[3] )


# MI model - vg prior
draws %>% 
  ggplot(aes(x=rr, y=0)) + 
  stat_halfeye(.width = c(0.025, 0.975)) + 
  scale_y_continuous(NULL, breaks = NULL) + 
  xlab("MI RR (ticagrelor vs. clopidogrel[ref])") +
  theme(panel.grid = element_blank()) +
  geom_vline(xintercept = c(0.9,1.11), colour="black", linetype="dashed") +
  geom_vline(xintercept = 1, colour="black", linetype="solid") +
  scale_x_continuous(limits = c(0, 2), expand = c(0, 0)) +
  ggtitle("MI with vague prior")



plot(tc4_mi_vg_identity)
grDevices::cairo_pdf("output/MI_vgI.pdf", width = 6, height = 6, family = "Arial")
plot(tc4_mi_vg_identity, digits = 2, base_size = 10)
dev.off()

prop0.9_mi <- sum(draws$rr<0.9)/nrow(draws)
prop1.0_mi <- sum(draws$rr<1.0)/nrow(draws)
prop1.1_mi <- sum(draws$rr>1.1)/nrow(draws)

cat("Pr(HR<0.9): ", prop0.9_mi,
    "\nPr(HR<1.0): ", prop1.0_mi,
    "\nPr(HR>1.1): ", prop1.1_mi,
    "\nPr(HR[0.9, 1.1]): ", 1-prop1.1_mi-prop0.9_mi)

```

Using a vague prior , the  MIposterior distribution is a N(`r temp[2]`, `r (temp[3] - temp[2])/1.96 `)       
Essentially same results from logit or identity link, as expected

Diagnostics
```{r}
pairs(tc4_mi_vg)
plot(tc4_mi_vg, ask=F)
```

```{r}
mcmc_areas(draws,
           pars = "rr",
           prob = 0.95)
```

## NA prior - with identity link
### Canada & US MI data from PLATO

```{r fig.width=10, fig.height=6, warning=FALSE, message=FALSE}

# NA total Canada 35/31 US 178/142 deaths CDN 7/5  US 29/24
# assume no strokes so MI CDN 28/26 US 149/118

### MI

plot_NA_MI <- meta::metabin(event.e=c(149,28), event.c=c(118,26), n.e=c(706,200), n.c=c(706,200), sm="RR", method ="I", 
    studlab=c("US", "CDN"), prediction=TRUE, comb.random =TRUE) #, comb.random =TRUE
meta::forest(plot_NA_MI, rows.gr=50)

pdf("PI_death_NA.pdf", height = 12, width = 11)
meta::forest(plot_NA_MI, rows.gr=50)
dev.off()

```

Use this data to create priors for effectiveness of ticagrelor vs. clopidogrel 

```{r warning=FALSE, message=FALSE, cache=TRUE}
# Use stan directly to get the prior for each arm
basic_binom <- cmdstan_model("binomial1.stan")

set.seed(123)
tic <- rbinom(906, size=1, prob= 0.195)
clop <- rbinom(906, size=1, prob= 0.159)

basic_binom_fit <- basic_binom$sample(data=list(N = length(tic),
                                                x = tic),
                                      iter_warmup=1000,
                                      iter_sampling=1000,
                                      refresh = 0,
                                      chains = 1)
print(basic_binom_fit)

basic_binom_fit1 <- basic_binom$sample(data=list(N = length(clop),
                                                x = clop),
                                      iter_warmup=1000,
                                      iter_sampling=1000,
                                      refresh = 0,
                                      chains = 1)

print(basic_binom_fit1)

RR <- as_draws_df(basic_binom_fit)[,2] / as_draws_df(basic_binom_fit1)[,2]
NA_prior <- quantile(RR[,1], c(0.025, 0.5, 0.975))
paste0("NA prior for MI is mean = ", round(NA_prior[2],2), " with 95% CI ", round(NA_prior[1], 2)," to ", round(NA_prior[3],2), " or n(", round(NA_prior[2], 2),", ", round((NA_prior[3]-NA_prior[2])/1.962,2), ")")
```


```{r cache=TRUE, warning=FALSE, message=FALSE}
b_prior_NA <- c(prior(normal(.16,.1), class=b, coef=drug0),
              prior(normal(.195,.1), class=b, coef=drug1))

tc4_mi_NA_identity <- brm(
  formula = f,
  data = TC4_mi,
  family = binomial(link = "identity"), #identity link probably better than get absolute numbers out can transform post
  chains = 4, warmup = 1000, iter = 2000, seed = 123,
  refresh = 0,
  prior=b_prior_NA, #prior(student_t(3, log(1), 5 ), class=b, coef=dapt_cat1)
  file = "tc4_mi_NA",
  refresh = 0
)


prior_summary(tc4_mi_NA_identity)
print(tc4_mi_NA_identity) # equivalent summary(m_logit)

draws <- as_draws_df(tc4_mi_NA_identity) # posterior_samples is equivalent command
# what have we done?
head(draws)

draws <- draws %>% 
  # rename and drop the unneeded columns
  transmute(p0 = b_drug0,   # expit not used since identity link
            p1 = b_drug1) %>% 
  # compute the OR
  mutate(or = (p1 / (1 - p1)) / (p0 / (1 - p0)), rr = p1 /p0)
head(draws)

temp <- round(quantile(draws$rr, c(0.025, 0.5, 0.975)),2)
paste0("RR = ", temp[2],  " 95% CI ",  temp[1], " - ", temp[3] )


# MI model - NA prior
draws %>% 
  ggplot(aes(x=rr, y=0)) + 
  stat_halfeye(.width = c(0.025, 0.975)) + 
  scale_y_continuous(NULL, breaks = NULL) + 
  xlab("MI RR (ticagrelor vs. clopidogrel[ref])") +
  theme(panel.grid = element_blank()) +
  geom_vline(xintercept = c(0.9,1.11), colour="black", linetype="dashed") +
  geom_vline(xintercept = 1, colour="black", linetype="solid") +
  scale_x_continuous(limits = c(0, 2), expand = c(0, 0)) +
  ggtitle("NA prior")


plot(tc4_mi_NA_identity)
grDevices::cairo_pdf("output/MI_NAprior.pdf", width = 6, height = 6, family = "Arial")
plot(tc4_mi_NA_identity, digits = 2, base_size = 10)
dev.off()

prop0.9_mi <- sum(draws$rr<0.9)/nrow(draws)
prop1.0_mi <- sum(draws$rr<1.0)/nrow(draws)
prop1.1_mi <- sum(draws$rr>1.1)/nrow(draws)

cat("Pr(HR<0.9): ", prop0.9_mi,
    "\nPr(HR<1.0): ", prop1.0_mi,
    "\nPr(HR>1.1): ", prop1.1_mi,
    "\nPr(HR[0.9, 1.1]): ", 1-prop1.1_mi-prop0.9_mi)

```

Using a NA prior from PLATO, the  MI posterior distribution is a N(`r temp[2]`, `r (temp[3] - temp[2])/1.96 `)       

Diagnostics
```{r}
pairs(tc4_mi_NA_identity)
plot(tc4_mi_NA_identity, ask=F)
```

```{r}
mcmc_areas(draws,
           pars = "rr",
           prob = 0.95)
```


## PLATO MI prior - with identity link


Use this NEJM data to create priors for effectiveness of ticagrelor  (504/9333) vs. clopidogrel (593/9291)

```{r warning=FALSE, message=FALSE, cache=TRUE}
# Use stan directly to get the prior for each arm
basic_binom <- cmdstan_model("binomial1.stan")

set.seed(123)
tic <- rbinom(9333, size=1, prob= 0.054) # 504/9333
clop <- rbinom(9291, size=1, prob= 0.0638) # 593/9291

basic_binom_fit <- basic_binom$sample(data=list(N = length(tic),
                                                x = tic),
                                      iter_warmup=1000,
                                      iter_sampling=1000,
                                      refresh = 0,
                                      chains = 1)
print(basic_binom_fit)

basic_binom_fit1 <- basic_binom$sample(data=list(N = length(clop),
                                                 x = clop),
                                       iter_warmup=1000,
                                       iter_sampling=1000,
                                       refresh = 0,
                                       chains = 1)

print(basic_binom_fit1)

RR <- as_draws_df(basic_binom_fit)[,2] / as_draws_df(basic_binom_fit1)[,2]
NA_prior <- quantile(RR[,1], c(0.025, 0.5, 0.975))
paste0("PLATO prior for MI is mean = ", round(NA_prior[2],2), " with 95% CI ", round(NA_prior[1], 2)," to ", round(NA_prior[3],2), " or n(", round(NA_prior[2], 2),", ", round((NA_prior[3]-NA_prior[2])/1.962,2), ")")
```

However this assumes completely pooled data with no heterogeneity between countries. Based on MACE and deaths where we have country specific data, we can see that there is some heterogeneity between countries with an approximate doubling of the SE.    
Therefore will use N(0.83, 0.12) as the prior for the PLATO data

```{r cache=TRUE, warning=FALSE, message=FALSE}
f = bf(y | trials(N) ~ 0 + drug)

m_PLATO <- brm(
  formula = f,
  data = TC4_mi,
  family = binomial(link = "identity"),
  prior = c(prior(beta(593,8698), class = b, coef = "drug0"),
            prior(beta(504,8294),class = b, coef = "drug1")),
  chains = 4, warmup = 1000, iter = 2000, seed = 123,
  refresh = 0
)

print(m_PLATO)

prior_summary(m_PLATO)
print(m_PLATO) # equivalent summary(m_logit)

draws <- as_draws_df(m_PLATO) # posterior_samples is equivalent command
# what have we done?
head(draws)

draws <- draws %>% 
  # rename and drop the unneeded columns
  transmute(p0 = b_drug0,   # expit not used since identity link
            p1 = b_drug1) %>% 
  # compute the OR
  mutate(or = (p1 / (1 - p1)) / (p0 / (1 - p0)), rr = p1 /p0)
head(draws)

temp <- round(quantile(draws$rr, c(0.025, 0.5, 0.975)),2)
paste0("RR = ", temp[2],  " 95% CI ",  temp[1], " - ", temp[3] )


# MI model - PLATO prior
draws %>% 
  ggplot(aes(x=rr, y=0)) + 
  stat_halfeye(.width = c(0.025, 0.975)) + 
  scale_y_continuous(NULL, breaks = NULL) + 
  xlab("MI RR (ticagrelor vs. clopidogrel[ref])") +
  theme(panel.grid = element_blank()) +
  geom_vline(xintercept = c(0.9,1.11), colour="black", linetype="dashed") +
  geom_vline(xintercept = 1, colour="black", linetype="solid") +
  scale_x_continuous(limits = c(0.5, 1.5), expand = c(0, 0)) +
  ggtitle("PLATO prior")


plot(m_PLATO)
grDevices::cairo_pdf("output/MI_PLATOprior.pdf", width = 6, height = 6, family = "Arial")
plot(m_PLATO, digits = 2, base_size = 10)
dev.off()

prop0.9_mi <- sum(draws$rr<0.9)/nrow(draws)
prop1.0_mi <- sum(draws$rr<1.0)/nrow(draws)
prop1.1_mi <- sum(draws$rr>1.1)/nrow(draws)

cat("Pr(HR<0.9): ", prop0.9_mi,
    "\nPr(HR<1.0): ", prop1.0_mi,
    "\nPr(HR>1.1): ", prop1.1_mi,
    "\nPr(HR[0.9, 1.1]): ", 1-prop1.1_mi-prop0.9_mi)
```


## BNMA MI prior - with identity link


Use this BNMA data to create priors for effectiveness of ticagrelor  (750 / 16020) vs. clopidogrel (1798/ 26189), which is  N(0.86, 0.13) as the prior for the BNMA data

```{r cache=TRUE, warning=FALSE, message=FALSE}
# useful function to get beta parameters from a normal distribution


f = bf(y | trials(N) ~ 0 + drug)

m_BNMA <- brm(
  formula = f,
  data = TC4_mi,
  family = binomial(link = "identity"),
  prior = c(prior(normal(1,.11), class = b, coef = "drug0"),
            prior(normal(0.96,0.11),class = b, coef = "drug1")),
  chains = 4, warmup = 1000, iter = 2000, seed = 123,
  refresh = 0
)

print(m_BNMA)

prior_summary(m_BNMA)
print(m_BNMA) # equivalent summary(m_logit)

draws <- as_draws_df(m_BNMA) # posterior_samples is equivalent command
# what have we done?
head(draws)

draws <- draws %>% 
  # rename and drop the unneeded columns
  transmute(p0 = b_drug0,   # expit not used since identity link
            p1 = b_drug1) %>% 
  # compute the OR
  mutate(or = (p1 / (1 - p1)) / (p0 / (1 - p0)), rr = p1 /p0)
head(draws)

temp <- round(quantile(draws$rr, c(0.025, 0.5, 0.975)),2)
paste0("RR = ", temp[2],  " 95% CI ",  temp[1], " - ", temp[3] )


# MI model - BNMA prior
draws %>% 
  ggplot(aes(x=rr, y=0)) + 
  stat_halfeye(.width = c(0.025, 0.975)) + 
  scale_y_continuous(NULL, breaks = NULL) + 
  xlab("MI RR (ticagrelor vs. clopidogrel[ref])") +
  theme(panel.grid = element_blank()) +
  geom_vline(xintercept = c(0.9,1.11), colour="black", linetype="dashed") +
  geom_vline(xintercept = 1, colour="black", linetype="solid") +
  scale_x_continuous(limits = c(0, 2), expand = c(0, 0)) +
  ggtitle("BNMA prior")


plot(m_BNMA)
grDevices::cairo_pdf("output/MI_BNMAprior.pdf", width = 6, height = 6, family = "Arial")
plot(m_BNMA, digits = 2, base_size = 10)
dev.off()

prop0.9_mi <- sum(draws$rr<0.9)/nrow(draws)
prop1.0_mi <- sum(draws$rr<1.0)/nrow(draws)
prop1.1_mi <- sum(draws$rr>1.1)/nrow(draws)

cat("Pr(HR<0.9): ", prop0.9_mi,
    "\nPr(HR<1.0): ", prop1.0_mi,
    "\nPr(HR>1.1): ", prop1.1_mi,
    "\nPr(HR[0.9, 1.1]): ", 1-prop1.1_mi-prop0.9_mi)
```

